---
title: "ESM 204 HW 3: Distributional consequences of climate policy"
author: "Bri Baker and Anna Talken"
date: "Spring 2021"
output:
  html_document:
    theme: cosmo
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)

options(scipen=999)
```

## Data Setup
```{r data info}
gas <- read_csv(here("HW3_data.csv")) %>% 
  clean_names() %>% 
  mutate(q_agg_gal = q_low_gallons + q_high_gallons)

gas_price <- 3 # $/gal
scc_ton <- 51 # $/metric ton
ton <- 2000 # lb
scc_lb <- scc_ton/ton # find scc per lb

```


## Question 1
```{r}
emission_gal <- 19.6 # pounds of co2 in a gallon of gas

mec <- emission_gal*scc_lb # mec as product of emissions and scc

```

The marginal external cost is $`r mec`/gal.


## Question 2

```{r demand curve}
d_agg <- lm(price_dollars ~ q_agg_gal, data = gas) # linear model of aggregate

d_agg_slope <- d_agg$coefficients[[2]] # aggregate slope
d_agg_intercept <- d_agg$coefficients[[1]] # aggregate intercept



# demand calculation functions take p or q input and use model outputs 

demand_q <- function(q, model){
  p <- model$coefficients[[1]] + model$coefficients[[2]]*q
return(p)
}

demand_p <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  return(q)
}

```

Aggregate demand: P = `r d_agg_slope`Q + `r d_agg_intercept`

```{r supply curve}

s_slope <- (gas_price/demand_p(gas_price, d_agg)) # calculate slope of mpc

```

```{r}
random_x <- data_frame("random" = 0:900000)
```

```{r}
# Plotting demand and supply curves

price_fitted_demand <- d_agg_intercept + d_agg_slope*gas$q_agg_gal #creates estimated line for demand

price_fitted_supply <- s_slope*gas$q_agg_gal #creates estimated line for supply

#aggregate_demand <- ggplot(data=gas, aes(x=q_agg_gal,y=price_dollars))+
  #geom_point() +
  #theme_minimal()+
  #labs(x = "Quantity of gas demanded (gal)",
      # y = "Price ($/gal)",
      # title = "Aggregate demand for gasoline using linear regression") +
  #geom_line(aes(y=price_fitted_demand, x=random_x$random)) +
 # geom_line(aes(y=price_fitted_supply, x=random_x$random)) #plotted the fitted relationship on the graph

#aggregate_demand

```

MPC: P = `r s_slope`Q

```{r surplus}

# surplus calculations take p or q input and find area based on model inputs

surplus_p <- function(p, model){
  q <- demand_p(p, model)
  cs <- 0.5*(model$coefficients[[1]] - p)*q
  return(cs)
}

surplus_q <- function(q, model){
  p <- demand_q(q, model)
  cs <- 0.5*(model$coefficients[[1]] - p)*q
  return(cs)
}


# use function

cs_initial <- surplus_p(gas_price, d_agg) # aggregate CS

ps_initial <- 0.5*(gas_price*demand_p(gas_price, d_agg)) # PS

```

Consumer benefit: $`r round(cs_initial, 2)`

Producer benefit: $`r round(ps_initial, 2)`

```{r env cost initial}

env_costs <- demand_p(gas_price, d_agg)*mec # cost as product of Q and MEC

```

Environmental cost: $`r round(env_costs, 2)`


## Question 3

```{r demands}

d_low <- lm(price_dollars ~ q_low_gallons, data = gas) # create linear model for low
d_high <- lm(price_dollars ~ q_high_gallons, data = gas) # create linear model for high

cs_low_initial <- surplus_p(gas_price, d_low) # CS of low at P = $3

cs_high_initial <- surplus_p(gas_price, d_high) # CS of low at P = $3

```

High income consumer benefit: $`r round(cs_high_initial, 2)`

Low income consumer benefit: $`r round(cs_low_initial, 2)`


## Question 4

```{r tax}

tax <- mec # since mec curve is constant, optimal tax = mec

```

The optimal tax is $`r round(tax, 2)`.

### a. The amount of gasoline produced and consumed:
```{r q_tax}

# function that represents q when MSC = Aggregate demand
equi_q <- function(q){
   d_agg_slope * q - s_slope * q + d_agg_intercept - mec
}

q_range <- c(0, 800000)

# solve function
q_tax <- uniroot(equi_q, q_range)

q_tax <- q_tax$root # pull Q when MSC = Aggregate demand

```

`r q_tax` gallons

### b. The price of gasoline:
```{r p_tax}

p_tax <- demand_q(q_tax, d_agg) # calculate P when MSC = Aggregate demand

```

The price of gasoline changes from $`r gas_price` to $`r round(p_tax, 2)`.


### c. Overall welfare of "high" income consumers: 
```{r welfare delta high}

cs_high_tax <- surplus_p(p_tax, d_high) # cs at taxed equilibrium

high_loss_tax <- cs_high_initial-cs_high_tax # delta

```

New high consumer welfare is $`r round(cs_high_tax, 2)` with a net loss of $`r round(high_loss_tax, 2)`.


### d. Overall welfare of "low" income consumers:
```{r welfare delta low}

env_cost_tax <- mec * q_tax # env burden when taxed

# cs at taxed equilibrium when bear burden env cost
cs_low_tax <- surplus_p(p_tax, d_low)-env_cost_tax

low_loss_tax <- cs_low_initial-cs_low_tax # delta

```
New low consumer welfare is $`r round(cs_low_tax, 2)` with a net loss of $`r round(low_loss_tax, 2)`.


### e. Gas producers:
```{r}

ps_tax <- 0.5*(p_tax-tax)*q_tax # ps at taxed equilibrium
ps_loss_tax <- ps_initial-ps_tax # delta

```
New producer surplus is `r round(ps_tax, 2)` with a net loss of $`r round(ps_loss_tax, 2)`


### f. Total environmental damage: 
$`r round(env_cost_tax, 2)`

### g. Total tax revenue generated: 
$`r round(env_cost_tax, 2)`



## Question 5

```{r range scc}

range_benefits <- tribble(~scc, ~mec, ~cb_high, ~cb_low, ~ps) # initialize dataframe
                          
scc_range <- c(51, 75, 100, 125, 150) # range of values of interest

high_per <- demand_p(3, d_high)/demand_p(3, d_agg) # percent of tax rev to high
low_per <- demand_p(3, d_low)/demand_p(3, d_agg) # percent of tax rev to low


for (n in scc_range){
  scc <- n # set scc as index in list
  mec_1 <- 19.6*scc/2000 # find mec at given scc
  tax <- mec_1 # find tax per unit
 
  q_tax <- uniroot(equi_q, q_range) # solve for Q at this scc
  
  q_tax <- q_tax$root # pull Q
  
  p_tax <- demand_q(q_tax, d_agg) # solve for P at this scc
  
  tax_rev <- tax * q_tax # find tax rev at this SCC
  
  cb_high <- surplus_q(q_tax, d_high) + high_per * tax_rev # calculate CB as CS + redistributed tax
  
  cb_low <- surplus_q(q_tax, d_low) + low_per * tax_rev # calculate CB as CS + redistributed tax
  
  ps <- ps_tax <- 0.5*(p_tax-tax)*q_tax # calculate PS
  
  sltn <- c(scc, mec_1, cb_high, cb_low, ps) # make vector of solutions
  
  range_benefits <- rbind(range_benefits, sltn) # add to df
}
  
# lol rename bc i forgot the right way to bind to a tibble

range_benefits_table <- range_benefits %>% 
  rename(
    "Social Cost of Carbon" = "X51",
    "Marginal External Cost" = "X0.4998",
    "Total High Consumer Benefit" = "X4630436.03078725",
    "Total Low Consumer Benefit" = "X10725144.9471771",
    "Gas Producer Benefit" = "X823485.334886087"
  )
range_benefits_table %>% 
  kbl(caption ="Table 1. Costs and Benefits in dollars of various CSS") %>% 
  kable_paper("hover", full_width = T)

```


## Question 6


### a. Total gas consumption
```{r electric}
gas <- gas %>% 
  mutate(
    electric_high = q_high_gallons/2, # electric reduces high consumer demand by half
    electric_agg = electric_high + q_low_gallons # new aggregate
  )

d_elec <- lm(price_dollars ~ electric_agg, data = gas) # linear model of aggregate

d_elec_slope <- d_elec$coefficients[[2]] # pull slope
d_elec_intercept <- d_elec$coefficients[[1]] # pull intercept


# function to find q when MSC = Aggregate demand
equi_ele_q <- function(q){
   d_elec_slope * q - s_slope * q + d_elec_intercept - mec
}

# Solve
q_ele <- uniroot(equi_ele_q, q_range)

q_ele <- q_ele$root # pull Q
```
`r q_ele`


### b. Total environmental externality
```{r env cost ele}

env_cost_ele <- mec * q_ele # environmental cost when electric

```
$`r round(env_cost_ele, 2)`


### c. Tax for same damage
```{r}
analagous_tax <- demand_q(q_ele, d_agg) - gas_price # tax for same env cost
```
$`r round(analagous_tax, 2)`


