---
title: "ESM 204 HW 3: Distributional consequences of climate policy"
author: "Bri Baker and Anna Talken"
date: "Spring 2021"
output:
  html_document:
    theme: cosmo
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)

options(scipen=999)
```

## Data Setup
```{r data info}
gas <- read_csv(here("HW3_data.csv")) %>% 
  clean_names() %>% 
  mutate(q_agg_gal = q_low_gallons + q_high_gallons)

gas_price <- 3 # $/gal
scc_ton <- 51 # $/metric ton
ton <- 2000 # lb
scc_lb <- scc_ton/ton

```


## Question 1
```{r}
emission_gal <- 19.6 # pounds of co2 in a gallon of gas

mec <- emission_gal*scc_lb

```

The marginal external cost is $`r mec`/gal.


## Question 2

```{r demand curve}
d_agg <- lm(price_dollars ~ q_agg_gal, data = gas) #create linear model

d_agg_slope <- d_agg$coefficients[[2]] #finds the slope
d_agg_intercept <- d_agg$coefficients[[1]] #finds the intercept

demand_q <- function(q, model){
  p <- model$coefficients[[1]] + model$coefficients[[2]]*q
return(p)
}

demand_p <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  return(q)
}

```

Aggregate demand: p = `r d_agg_slope`q + `r d_agg_intercept`

```{r supply curve}
s_slope <- (3/demand_p(3,d_agg))

s_intercept <- 0
```

```{r}
random_x <- data_frame("random" = 0:900000)
```

```{r}
# Plotting demand and supply curves

price_fitted_demand <- d_agg_intercept + d_agg_slope*gas$q_agg_gal #creates estimated line for demand

price_fitted_supply <- s_intercept + s_slope*gas$q_agg_gal #creates estimated line for supply

#aggregate_demand <- ggplot(data=gas, aes(x=q_agg_gal,y=price_dollars))+
  #geom_point() +
  #theme_minimal()+
  #labs(x = "Quantity of gas demanded (gal)",
      # y = "Price ($/gal)",
      # title = "Aggregate demand for gasoline using linear regression") +
  #geom_line(aes(y=price_fitted_demand, x=random_x$random)) +
 # geom_line(aes(y=price_fitted_supply, x=random_x$random)) #plotted the fitted relationship on the graph

#aggregate_demand

```

Supply: p = `r s_slope`q

```{r surplus}

surplus_p <- function(p, model){
  q <- demand_p(p, model)
  cs <- 0.5*(model$coefficients[[1]] - p)*q
  return(cs)
}

surplus_q <- function(q, model){
  p <- demand_q(q, model)
  cs <- 0.5*(model$coefficients[[1]] - p)*q
  return(cs)
}

cs_initial <- surplus_p(3, d_agg)

ps_initial <- 0.5*(3*demand_p(3, d_agg))

```

Consumer benefit: $`r cs_initial`

Producer benefit: $`r ps_initial`

```{r env cost initial}
env_costs <- demand_p(3, d_agg)*mec
```

Environmental cost: $`r env_costs`


## Question 3

```{r demands}

d_low <- lm(price_dollars ~ q_low_gallons, data = gas) # create linear model for low
d_high <- lm(price_dollars ~ q_high_gallons, data = gas) # create linear model for high

cs_low_initial <- surplus_p(3, d_low)

cs_high_initial <- surplus_p(3, d_high)

```

High income consumer benefit: $`r cs_high_initial`

Low income consumer benefit: $`r cs_low_initial`


## Question 4

```{r tax}

tax <- mec # since mec curve is constant, optimal tax = mec

```

The optimal tax is $`r tax`.

(a) The amount of gasoline produced and consumed:
```{r q_tax}
equi_q <- function(q){
   d_agg_slope * q - s_slope * q + d_agg_intercept - mec
}

q_range <- c(0, 800000)

q_tax <- uniroot(equi_q, q_range)

q_tax <- q_tax$root

```

`r q_tax` gallons

(b) The price of gasoline:
```{r p_tax}

p_tax <- demand_q(q_tax, d_agg)

```

$`r round(p_tax, 2)`


(c) Overall welfare of "high" income consumers: 
```{r}
cs_high_tax <- surplus_p(p_tax, d_high)

high_loss_tax <- cs_high_initial-cs_high_tax

```
$`r high_loss_tax`


(d) Overall welfare of "low" income consumers:
```{r}
env_cost_tax <- mec*q_tax

cs_low_tax <- surplus_p(p_tax, d_low)-env_cost_tax

low_loss_tax <- cs_low_initial-cs_low_tax

```
$`r cs_low_tax`


(e) Gas producers:
```{r}

ps_tax <- 0.5*(p_tax-tax)*q_tax
ps_loss_tax <- ps_initial-ps_tax

```
Net loss: $`r ps_loss_tax`

(f) Total environmental damage: $`r round(env_cost_tax, 2)`
(g) Total tax revenue generated: $`r round(env_cost_tax, 2)`



## Question 5

```{r}
range_benefits <- tribble(~scc, ~mec, ~cb_high, ~cb_low, ~ps)
scc_range <- c(51, 75, 100, 125, 150)

high_per <- demand_p(3, d_high)/demand_p(3, d_agg)
low_per <- demand_p(3, d_low)/demand_p(3, d_agg)


for (n in scc_range){
  scc <- n
  mec <- 19.6*scc/2000
  tax <- mec
  
  q_range <- c(0, 800000)
  
  q_tax <- uniroot(equi_q, q_range)
  
  q_tax <- q_tax$root
  
  p_tax <- demand_q(q_tax, d_agg)
  
  tax_rev <- tax * q_tax
  
  cb_high <- surplus_q(q_tax, d_high) + high_per * tax_rev
  
  cb_low <- surplus_q(q_tax, d_low) + low_per * tax_rev
  
  ps <- ps_tax <- 0.5*(p_tax-tax)*q_tax
  
  sltn <- c(scc, mec, cb_high, cb_low, ps)
  
  range_benefits <- rbind(range_benefits, sltn) # need to figure this out???? why does it overwrite the names?
}
  

```

